Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-05-15 Manuscript/manuscript.Rmd
manuscript edits

Therefore, in this study, we aimed to develop an ensemble framework composed of different combinations of data driven characteristics used to address the issues presented above. We used our framework to systematically compare performances of random forest classifiers using a composite performance metric, MCC. Our goal was to determine a suitable combination of data re-sampling and feature engineering techniques that both corrected the inherent imbalanced classes and provided the best spatial representation linking genomic annotations to TAD boundaries. Additionally, we used our framework to directly compare the influence of Hi-C contact matrix resolution on model performances. We evaluated Hi-C data for both the GM12878 and K562 cell line obtained at 10 kb, 25 kb, 50 kb, and 100 kb resolution. For each resolution, we examined the inclusion of three different predictor types, including distance as the measure of association between genomic elements. We then tested four different re-sampling techniques aimed at creating balanced classes. Noninformative features were removed using a feature reduction algorithm and random forest classifications were built to assess model performance, and investigate which genomic annotations were most predictive of TAD boundaries for a given system from the complete ensemble.

@@

A regularization technique using a combination of the *l_{1}* and *l_{2}* penalties, known as the elastic-net, was used to reduce the feature space. The elastic-net was chosen in order to suitably reduce the feature space while also maintaining the inclusion of multiple correlated predictors. For each elastic-net regularization, 10-fold cross-validation was used to tune the parameters that controlled the penalty terms. The parameters that maximized the mean AUPRC across each cross-fold was chosen. All elastic-net algorithms were implemented in R using the *caret* package version 6.0 [@kuhn2012caret].

>>>>

Therefore, in this study, we aimed to develop an ensemble framework composed of different combinations of data driven characteristics used to address the issues presented above. We used our framework to systematically compare performances of random forest classifiers using a composite performance metric, MCC. Our goal was to determine a suitable combination of data re-sampling and feature engineering techniques that both corrected the inherent imbalanced classes and provided the best spatial representation linking genomic annotations to TAD boundaries. Additionally, we used our framework to directly compare the influence of Hi-C contact matrix resolution on model performances. We evaluated Hi-C data for both the GM12878 and K562 cell line obtained at 10 kb, 25 kb, 50 kb, and 100 kb resolution. For each resolution, we examined the inclusion of three different predictor types, including distance as the measure of association between genomic elements. We then tested four different re-sampling techniques aimed at creating balanced classes. This process was performed on chromosome-specific regions on the linear genome. Noninformative features were removed using a feature reduction algorithm and classifiers were built to assess model performance, and investigate which genomic annotations were most predictive of TAD boundaries for a given system from the complete ensemble.

@@

Architectural proteins tend to exhibit an extensive pattern of colocalization (correlation) [@van2014insulator]. To measure the pairwise correlation of each cell line specific annotation, we used the Jaccard Index. The Jaccard Index measures the ratio of the number of intersecting base pairs between two sets to the number of base pairs in the union, subtracted by the intersection [@favorov2012exploring]. The quantity ranges from 0.0 to 1.0, where 0.0 represents no overlap (no correlation) and 1.0 represent complete overlap (complete correlation).

In order to suitably reduce the feature space while also maintaining the inclusion of possibly colocalized annotations, we implemented an elastic-net regularization technique prior to modelling. The technique using a combination of the *l_{1}* and *l_{2}* penalties. For each elastic-net regularization, 10-fold cross-validation was used to tune the parameters that controlled the penalty terms. <!--The parameters that maximized the mean AUPRC across each cross-fold was chosen.--> All elastic-net algorithms were implemented in R using the *caret* package version 6.0 [@kuhn2012caret].
